Lifting CDCL to Relational Domains

### 2016-10-17 learning.tex
learning

For our example, there exists multiple incomaparable reasons for conflict,
marked as {\em cut0, cut1}, in figure~\ref{conflict}.  Here, cut0 is the first UIP.  
Choosing cut0 yields a learnt clause 
$L_0: (x+y<15 \vee -x-y<-15 \vee -x+y<5 \vee x-y<-5 \vee y-10<0 \vee -y+10<0)$, 
which is obtained by negating the reason for conflict.  
%
\begin{figure}
%\caption{An example of Octagon}\label{octagon}
\scalebox{.65}{\import{figures/}{conflict_graph.pspdftex}}
\label{conflict}
\end{figure} 
%    
\paragraph{\textbf{Learning in Relational Lattice}}
We model a learning as a meet irreducible deduction transformer, called 
{\em abstract unit transformer (AUnit)}.  For an 
abstract lattice $A$ with complementable meet irreducibles 
and a set of conflicting meet irreducibles (countermodels) 
$C \subseteq A$ of $\varphi$ such that \meet{C} does not 
satisfy $\varphi$, $AUnit_C: A \rightarrow A$ is defined as follows. 
\[ AUnit(a) =
 \begin{cases}
  \bot       & \quad \text{if } a \sqsubseteq \meet{C} \\
  \bar{t}    & \quad \text{if } t \in C \; \text{and} \; \forall t' \in C
  \setminus \{t\}. a  \sqsubseteq t' \\
  \bar{t}    & \quad \text{if } t \in C \; \text{and} \; \forall t' \in C \setminus \{t\}. a
  \not\sqsubseteq t', \; \text{then} \\ 
             & \quad \forall a' \subseteq decomp(a) \;  \text{and} \; \forall a'' \in a \setminus \{a'\}, \\ 
             & \quad \meet{a''} \sqsubseteq r' \; \text{and} \; a' \not\sqsubseteq t  \\
  \top & \quad \text{otherwise} \\
 \end{cases}
\]
For example, let $C = \langle x+z \geq 10,  k \geq 0 \rangle$ be the conflicting set 
of meet irreducibles of $\varphi$.  Let $a= \langle x+y \geq 2, y+z \geq 5, x+z \leq 15, k:[1,1] \rangle$, 
then $AUnit_C(a) = (x+z < 10)$.  Applying the third rule, we get a decomposition 
of $a$ into $a'$ and $a''$ where 
$a'= \langle x+y \geq 2, y+z \geq 5, x+z \leq 15 \rangle \not\sqsubseteq (x+z \geq 10)$, and 
$a''= \langle k:[1,1] \rangle \sqsubseteq (k \geq 0)$.  $AUnit_C(a)$ returns a new 
deduction $x+z \leq 9$, which constrains the bound of the octagonal constraint.  
Let us consider another example, where $C = \langle x \geq 2, x \leq 5, y \leq 7 \rangle$
and let $a = \langle x:[3,4], y:[5,6] \rangle$, then $AUnit_C(a) = \bot$ using
the first rule, since $a \sqsubseteq C$.  Now, let $a' = \langle x:[3,4] \rangle$, 
then $AUnit_C(a') = (y > 7)$ using second rule, since $a' \sqsubseteq (x \geq 2)$
and $a' \sqsubseteq (x \leq 5)$.

>>>>

For our example, there exists multiple incomaparable reasons for conflict,
marked as {\em cut0, cut1}, in figure~\ref{conflict}.  Here, cut0 is the first UIP.  
Choosing cut0 yields a learnt clause 
$L_0: (x+y<15 \vee -x-y<-15 \vee -x+y<5 \vee x-y<-5 \vee y-10<0 \vee -y+10<0)$, 
which is obtained by negating the reason for conflict.  
%
\begin{figure}
%\caption{An example of Octagon}\label{octagon}
\scalebox{.65}{\import{figures/}{conflict_graph.pspdftex}}
\label{conflict}
\end{figure} 
%    
\paragraph{\textbf{Learning in Relational Lattice}}
We model learning as a meet irreducible deduction transformer, called 
{\em abstract unit transformer (AUnit)}.  For an 
abstract lattice $A$ with complementable meet irreducibles 
and a set of conflicting meet irreducibles (countermodels) 
$C \subseteq A$ of $\varphi$ such that \meet{C} does not 
satisfy $\varphi$, $AUnit_C: A \rightarrow A$ is defined as follows. 
\[ AUnit(a) =
 \begin{cases}
  \bot       & \quad \text{if } a \sqsubseteq \meet{C} \\
  \bar{t}    & \quad \text{if } t \in C \; \text{and} \; \forall t' \in C
  \setminus \{t\}. a  \sqsubseteq t' \\
  \bar{t}    & \quad \text{if } t \in C \; \text{and} \; \forall t' \in C \setminus \{t\}. a
  \not\sqsubseteq t', \; \text{then} \\ 
             & \quad \forall a' \subseteq decomp(a) \;  \text{and} \; \forall a'' \in a \setminus \{a'\}, \\ 
             & \quad \meet{a''} \sqsubseteq r' \; \text{and} \; a' \not\sqsubseteq t  \\
  \top & \quad \text{otherwise} \\
 \end{cases}
\]
For example, let $C = \langle x+z \geq 10,  k \geq 0 \rangle$ be the conflicting set 
of meet irreducibles of $\varphi$ and let $a= \langle x+y \geq 2, y+z \geq 5,
x+z \leq 15, k:[1,1] \rangle$.  Applying the third rule of $AUnit_C$, we get a 
decomposition of $a$ into $a'$ and $a''$ where 
$a'= \langle x+y \geq 2, y+z \geq 5, x+z \leq 15 \rangle \not\sqsubseteq (x+z \geq 10)$, and 
$a''= \langle k:[1,1] \rangle \sqsubseteq (k \geq 0)$.  Then, $AUnit_C(a) = (x+z < 10)$, 
which constrains the bound of the octagonal constraint.  
Let us consider another example, where $C = \langle x \geq 2, x \leq 5, y \leq 7 \rangle$
and let $a = \langle x:[3,4], y:[5,6] \rangle$, then $AUnit_C(a) = \bot$ using
the first rule, since $a \sqsubseteq C$.  Now, let $a' = \langle x:[3,4] \rangle$, 
then $AUnit_C(a') = (y > 7)$ using second rule, since $a' \sqsubseteq (x \geq 2)$
and $a' \sqsubseteq (x \leq 5)$.
